---
title: "Microbiome Analysis Report"
author: "Priyadarshini K"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
params:
  patient_name: "P009712"
  output_dir: "./P009712"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 6,
  out.width = '80%'
)
library(readr)
library(readxl)
library(dplyr)
library(knitr)
library(data.table)
library(knitr)
library(kableExtra)
library(ggplot2)

# -----------------------------
# Parameters
# -----------------------------
patient_name <- params$patient_name
base_dir <- params$output_dir

# -----------------------------
# Helper functions
# -----------------------------
# Robust helper to load files
load_file <- function(patient_id, suffix) {
  # Top-level folder
  path1 <- file.path(base_dir, paste0(patient_id, suffix))
  # Nested folder
  path2 <- file.path(base_dir, patient_id, paste0(patient_id, suffix))

  if (file.exists(path1)) return(path1)
  if (file.exists(path2)) return(path2)
  # Return path1 anyway for warning
  return(path1)
}

# Safe plotting helper
safe_plot <- function(filepath) {
  if (file.exists(filepath)) {
    knitr::include_graphics(filepath)
  } else {
    cat("Image not available:", filepath, "\n")
  }
}
```


# Patient Information

**Patient ID:** `r params$patient_name`  
**Report Date:** `r Sys.Date()`  

\newpage

## fastq-screen decontamination QC R1
```{r fastq-screenr1, echo=FALSE}
safe_plot(load_file(patient_name, "_cleaned_paired_1_screen.png"))
```
## fastq-screen decontamination QC R2
```{r fastq-screenr2, echo=FALSE}
safe_plot(load_file(patient_name, "_cleaned_paired_2_screen.png"))
```

# Kraken2 Taxonomy Reports
```{r kraken, echo=FALSE}
kraken_file <- load_file(patient_name, ".kraken2.report")


if (file.exists(kraken_file)) {
  load_kraken_report <- function(filepath, sample_name) {
    df <- fread(filepath, header = FALSE, sep = "\t")
    colnames(df) <- c("Percent", "Reads_Clade", "Reads_Direct", "Rank", "TaxID", "Name")
    df$Sample <- sample_name
    return(df)
  }

  sample_df  <- load_kraken_report(kraken_file, "Patient")

  kable(head(sample_df, 10), caption = "Top 10 entries from patient Kraken2 report")
} else {
  cat("Kraken2 report not found for patient:", patient_name, "\n")
}
```
## Krona Plot
```{r krona, echo=FALSE}
safe_plot(load_file(patient_name, "_krona.png"))
```
\newpage

# Differential Abundance Analysis
```{r da, echo=FALSE}
da_file <- load_file(patient_name, "_Differentially_Abundant_Microbes.xlsx")
if (file.exists(da_file)) {
  da <- read_xlsx(da_file)
  kable(head(da, 10), caption = "Top Differentially Abundant Microbes")
} else {
  cat("No differential abundance results for patient:", patient_name, "\n")
}

```
## Heatmap
```{r volcano, echo=FALSE}
safe_plot(load_file(patient_name, "_heatmap_top30.png"))
```

### Top 10 Species in Patient Sample
```{r topspecies, echo=FALSE}
safe_plot(load_file(patient_name, "_Top10_Patient_Species.png"))
```

### Top 10 Species: Patient vs Control
```{r pc, echo=FALSE}
safe_plot(load_file(patient_name, "_Top10_Patient_vs_Control.png"))
```
\newpage

# Microbe Gene Rank Interaction Network Table
```{r rank, echo=FALSE}
rank_csv <- load_file(patient_name, "_Microbe_Gene_RankSimilarity.csv")
if (file.exists(rank_csv)) {
  rank_df <- read_csv(rank_csv, show_col_types = FALSE)
  rank_top <- rank_df %>% arrange(desc(RankSimilarity)) %>% head(10)
  kable(rank_top, caption = "Top 10 Microbe–Gene Rank Similarities")
} else {
  cat("No rank similarity file found for patient:", patient_name, "\n")
}

```
## Microbe Gene Rank Interaction Network
```{r rankplot, echo=FALSE}
safe_plot(load_file(patient_name, "_Microbe_Gene_rank_network.png"))
```
\newpage

# Diversity
## Alpha Diversity
```{r adiv, echo=FALSE}
safe_plot(load_file(patient_name, "_AlphaDiversity.png"))
```

## Beta Diversity
```{r bdiv, echo=FALSE}
safe_plot(load_file(patient_name, "_BetaDiversity.png"))
```
\newpage

# Microbe-Metabolite_Gene_Disease Interaction
```{r mmgd, echo=FALSE}
safe_plot(load_file(patient_name, "_top_microbe_metabolite_gene_disease_network.png"))
```

## Microbe-Metabolite_Gene_Disease Interaction Table
```{r mmgdplot, echo=FALSE}
gutm_csv <- load_file(patient_name, "_microbe_metabolite_gene_disease_network.csv")
if (file.exists(gutm_csv)) {
  gutm_df <- read_csv(gutm_csv, show_col_types = FALSE)
  gutm_top <- gutm_df %>% head(10)
  kable(gutm_top, caption = "Top 10 Microbe–Metabolite–Gene–Disease Interactions")
} else {
  cat("No MMGD interactions file found for patient:", patient_name, "\n")
}

```

